Model

### 2019-11-14 stats/model.Rmd
update model with prior estimation integrated

@@

var subjects = levels(priorData, "subid")

@@

  var subject_sigma = uniformDrift({a: 0, b:3, width: 0.2})

  var sampleSub = function(age){
    return gaussianDrift({ mu: age, sigma: subject_sigma, width: 0.1 })
  }

  foreach(function(subid){
      var subjectData = _.filter(priorData, {subid: subid})

@@

      var subj_age = subjectData[0].age

      var subjectPrior = sampleSub(subj_age)

      display(subjectPrior)

@@

  query.add(["intercept"], [prior_int])
  query.add(["slope"], [prior_slope])
  query.add(["sigma"], [subject_sigma])

@@

prior_data <- read_csv("../data/novelty.csv")%>%
  mutate(age = age_num - min(age_num))

prior_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, prior_params , sep='\n'),
  data = list(data = prior_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 1,
  cores = 1,
  inference_opts = list(method = "MCMC", samples = 1, burn = 1, verbose = T)
)

@@

@@

novel_bm <- brm(correct ~ age + (1 | subid), 
          data = prior_data, family = bernoulli(),
          control = list(adapt_delta = 0.95),
          sample_prior = F,
          save_all_pars = TRUE,
          iter = 20000)

@@

brm_plot <- tibble(
  age = prior_data$age,
  y = plogis(brm_params$Intercept + brm_params$age * age),
  model = "brm"
)

@@

brm_full_plot <- tibble(
  age = prior_data$age,
  y = plogis(brm_full_params$Intercept + brm_full_params$age_num * age),
  model = "brm_full"
)

# webppl model
webppl_plot <- tibble(
  age = prior_data$age,
  y = plogis(prior_summary%>%filter(Parameter == "intercept")%>%pull(param_map) + prior_summary%>%filter(Parameter == "slope")%>%pull(param_map) * age),
  model = "webppl"
)

@@

```{r work in progress}
sem_subj_str <- '
var data = dataFromR.data

@@

var familiars = levels(data, "familiar")
var subjects = levels(data, "subid")

@@

@@

@@

var all_log_likes = map(function(row) { 
  var prior = [row.prior, 1-row.prior]
  var itemLineParameters = all_item_parameters[row.item]
  var sem_knowledge = logistic(itemLineParameters.intercept + 
                               itemLineParameters.slope * row.age)

@@

  var pragModelPredictions = pragmaticListener({label: "novel_word"}, 
                                               prior, speakerOptimality, 
                                               sem_knowledge)

@@

  var globalPragModelPredictions = pragmaticListener({label: "novel_word"}, 
                                               prior, speakerOptimality, 
                                               global_sem_knowledge)

@@

  var priorOnlyModelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: prior}));
      return obj.shape == "novel_object" ? 1 : 0
       }})

@@

    query.add(["modelPrediction","pragmatic","noise", row.condition, row.familiarObject, row.subid], [row.age, Math.exp(noisyPragModelPredictions.score(1))])

@@

  query.add(["modelPrediction","globalPrag","free", row.condition, row.familiarObject, row.subid], [row.age, Math.exp(globalPragModelPredictions.score(1))])

  query.add(["modelPrediction","globalPrag","noise", row.condition, row.familiarObject, row.subid], [row.age, Math.exp(noisyGlobalPragModelPredictions.score(1))])

@@

Infer({model,
      samples: 50000,
      burn: 250000,
      verbose: T,
      lag: 10,
      verboseLag: 5000,
      method: "incrementalMH",
      onlyMAP: true,
      callbacks: [callback]});
'
```

@@

```{r}
sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, sem_subj_str , sep='\n'),
  data = list(data = me_model_data, comb_data = data_comb_log),
  data_var = "dataFromR",
  model_var = "model",
  chains = 2,
  cores = 2,
  packages = c("webppl_packages/webppl-sample-writer")
)

@@

model_pred <- sem_know_model %>%
  filter(a == "modelPrediction")%>%
  select(-a, -f)%>%
  rename(model = b, 
         parameter = c,
         alignment = d,
         item = e,
         age = g, 
         pred = h)

@@

```
## Plotting model parameters

@@

```{r}
data_comb_plot<-data_comb%>%
  mutate(model = "data")

@@

plot_model_pred_prag_free <- model_pred_prag_free%>%
  mutate(model = paste(model, parameter, sep = "_"))%>%
  group_by(model, alignment, item, age)%>%
  summarise(mean = estimate_mode(pred),
            ci_lower = hdi_lower(pred),
            ci_upper = hdi_upper(pred))

>>>>

prior_data <- read_csv("../data/novelty.csv")

prior_model_data <- read_csv("../data/novelty.csv")%>%
  mutate(age = age_num - min(age_num))

me_model_data <- read_csv("../data/me.csv")%>%
    filter(trial != "train1",
           trial != "train2")%>%
  mutate(age = age_num - min(age_num))

@@

var priorSubjects = levels(priorData, "subid")

@@

  var prior_subject_sigma = uniformDrift({a: 0, b:3, width: 0.2})

  var priorSampleSub = function(age){
    return gaussianDrift({ mu: age, sigma: prior_subject_sigma, width: 0.1 })
  }

  foreach(function(subid){
      var priorSubjectData = _.filter(priorData, {subid: subid})

@@

      var subj_age = priorSubjectData[0].age

      var subjectPrior = priorSampleSub(subj_age)

@@

    }, priorSubjectData)

  }, priorSubjects)

  query.add(["intercept"], [prior_int])
  query.add(["slope"], [prior_slope])
  query.add(["sigma"], [prior_subject_sigma])

@@

prior_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, prior_params , sep='\n'),
  data = list(data = prior_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 1,
  cores = 1,
  inference_opts = list(method = "MCMC", samples = 1, burn = 1, verbose = T)
)

@@

prior_model <- readRDS("../saves/prior_model.rds")

@@

novel_bm <- brm(correct ~ age + (1 | subid), 
          data = prior_model_data, family = bernoulli(),
          control = list(adapt_delta = 0.95),
          sample_prior = F,
          save_all_pars = TRUE,
          iter = 20000)

@@

brm_plot <- tibble(
  age = prior_model_data$age,
  y = plogis(brm_params$Intercept + brm_params$age * age),
  model = "brm"
)

@@

brm_full_plot <- tibble(
  age = prior_model_data$age,
  y = plogis(brm_full_params$Intercept + brm_full_params$age_num * age),
  model = "brm_full"
)

# webppl model
webppl_plot <- tibble(
  age = prior_model_data$age,
  y = plogis(prior_summary%>%filter(Parameter == "intercept")%>%pull(param_map) + prior_summary%>%filter(Parameter == "slope")%>%pull(param_map) * age),
  model = "webppl"
)

@@

### Combination Model

```{r work in progress}
sem_subj_str <- '

var data = dataFromR.data
var priorData = dataFromR.data

@@

var familiars = levels(data, "familiar")
var subjects = levels(data, "subid")
var priorSubjects = levels(priorData, "subid")

@@

//////////////// Prior ////////////////////////

  var prior_subject_sigma = uniformDrift({a: 0, b:3, width: 0.2})

  var priorSampleSub = function(age){
    return gaussianDrift({ mu: age, sigma: prior_subject_sigma, width: 0.1 })
  }

  foreach(function(subid){
      var priorSubjectData = _.filter(priorData, {subid: subid})

      var subj_age = priorSubjectData[0].age

      var subjectPrior = priorSampleSub(subj_age)

    }, priorSubjectData)

  }, priorSubjects)

  query.add(["parameter","parameters", "prior"], [prior_int, prior_slope])
  query.add(["parameter","sigma", "prior"], ["NA", prior_subject_sigma])

//////////////// Semantic knowledge and speaker optimality ////////////////////////

@@

//////////////// Model predictions and combination ////////////////////////

@@

var all_log_likes = map(function(row) {

  var priorReg = logistic(prior_int + prior_slope * row.age)

  var cond = row.condition

  var prior = function(cnd){
    var cndPrior = cnd == "congruent" ? [priorReg, 1-priorReg] : [1-priorReg, priorReg] ;
    return cndPrior
  }

  var priorComb = prior(cond)

  var itemLineParameters = all_item_parameters[row.item]

  var sem_knowledge = logistic(itemLineParameters.intercept + 
                               itemLineParameters.slope * row.age)

@@

  var pragModelPredictions = pragmaticListener({label: "novel_word"}, 
                                               priorComb, speakerOptimality, 
                                               sem_knowledge)

@@

  var globalPragModelPredictions = pragmaticListener({label: "novel_word"}, 
                                               priorComb, speakerOptimality, 
                                               global_sem_knowledge)

@@

  var priorOnlyModelPredictions = Infer({method: "enumerate", model: function(){
      var obj = sample( Categorical({vs: all_objects, ps: priorComb}));
      return obj.shape == "novel_object" ? 1 : 0
       }})

@@

@@

//  query.add(["modelPrediction","globalPrag","free", row.condition, row.familiarObject, row.subid], [row.age, Math.exp(globalPragModelPredictions.score(1))])

@@

Infer({model,
      samples: 50000,
      burn: 50000,
      verbose: T,
      lag: 1,
      verboseLag: 1000,
      method: "incrementalMH",
      onlyMAP: true,
      callbacks: [callback]});
'
```

@@

```{r}
sem_know_model<- webppl(
  program_code = paste(rsaUtils, rsaModel, sem_subj_str , sep='\n'),
  data = list(data = me_model_data, comb_data = data_comb_log, prior_data = prior_model_data),
  data_var = "dataFromR",
  model_var = "model",
  chains = 2,
  cores = 2,
  packages = c("webppl_packages/webppl-sample-writer")
)

@@

model_pred <- read_csv("../saves/model/samples78.csv") %>%
  filter(a == "modelPrediction")%>%
  select(-a)%>%
  rename(model = b, 
         parameter = c,
         alignment = d,
         id = f,
         item = e,
         age = g, 
         pred = h)

@@

model_pred %>%
  filter(item == "pawn")%>%
  arrange(age)
```
## Plotting model parameters

@@

@@

plot_model_pred_prag_free <- model_pred%>%
  mutate(model = paste(model, parameter, sep = "_"))%>%
  group_by(model, alignment, item, age)%>%
  summarise(mean = estimate_mode(pred),
            ci_lower = hdi_lower(pred),
            ci_upper = hdi_upper(pred))
